---
title: "Example HPC workflow and troubleshooting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example workflow and troubleshooting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script src="https://cdn.jsdelivr.net/npm/mermaid/dist/mermaid.min.js"></script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette walks you through a variety of different installation options and workflows for
analysing data in bulk on a high performance computing (HPC) platform.

## Loading package onto an "airgapped" HPC system

If working on a secure or "airgapped" system without internet access, using the R `remotes` package to
install `nutrientprofiler` will not work.

Instead, if there is a file/data transfer process available, the zipped package can be
transferred to the secure computational platform and installed locally. 

This process can be broken down into 3 steps:

1. Downloading the package/zipped archive to your local system
2. Transferring the files to the secure system (after any required security checks)
3. Installing without internet access on the secure system.

The instructions here assume access to some form of archived CRAN mirror so that other commonly used
packages are available. Please check through the packages listed in the code snippets and ensure you have
access to these on your system.

### 1. Downloading the package to your local machine

You can download a specific tagged release (see the 
[GitHub releases](https://github.com/Leeds-CDRC/nutrientprofiler/releases) page) or a development
pre-release version of the code from a specific branch of the repository.

Tagged releases allow for better reproducibility; however, if a pre-release version needs to be used,
reproducibility can still be ensured by recording the download link used and the access date.


#### 1.1 Downloading a tagged release

**The recommended method is to download the most recent tagged release as a `.tar.gz` archive.**

From your local desktop you can run the following R snippet to download the `v1.0.0` release of the package:

```R
# download the v1.0.0 release as a .tar.gz archive
# to your current directory
download.file("https://github.com/Leeds-CDRC/nutrientprofiler/archive/refs/tags/v1.0.0.tar.gz",
              dest = "./nutrientprofiler-v1.0.0.tar.gz")
```

Alternatively, this can be downloaded directly from the
[GitHub releases](https://github.com/Leeds-CDRC/nutrientprofiler/releases) page.

A different release can be specified by changing the release tag. Make sure to also update
the destination filename to prevent confusion:

```R
# replace <tag> with the required version number
download.file("https://github.com/Leeds-CDRC/nutrientprofiler/archive/refs/tags/<tag>.tar.gz",
              dest = "./nutrientprofiler-<tag>.tar.gz")
```

If required, zip file archives are also available on the 
[GitHub releases](https://github.com/Leeds-CDRC/nutrientprofiler/releases) page or can be downloaded
using the following R script below. Please note that these `.zip` archives require a slightly different
installation to `.tar.gz`; please read through the isntallation steps first.

```R
# download the <tag> release as a zip archive
# to your current directory
download.file("https://github.com/Leeds-CDRC/nutrientprofiler/archive/refs/tags/<tag>.zip",
              dest = "./nutrientprofiler-<tag>.zip")
```

#### 1.2 Downloading the current development version or a specific branch

If you want to install a pre-release version of a specific version on a branch, replace `tags/version-number`
in the url with `heads/branch-name` (and rename the destination file something sensible).

**Please record the date of download and most recent commit identifier as the branch may be updated or changed following your download and installation.**

For example, the following downloads the current package on the "VarEdits" branch of the repository:

```R
# replace "tags/version-number" with "heads/branch-name":
# in this example, we replaced "tags/v1.0.0" with "heads/VarEdits"
download.file("https://github.com/Leeds-CDRC/nutrientprofiler/archive/refs/heads/VarEdits.tar.gz",
              dest = "./nutrientprofiler-VarEdits.tar.gz")
```

Again, a zip file version can be downloaded by replacing `tar.gz` in the snippet above with `zip`. A zip
file can also be downloaded from the project [GitHub page](https://github.com/Leeds-CDRC/nutrientprofiler)
by navigating to the required branch and then using the green "Code" button to open the "Clone" option menu,
and selecting the "Download zip" option.

### 2. Transferring your code to the secure platform

This step will vary depending on the data transfer policies and process enforced by your institution. While
the package archive can be unzipped for testing, it should be saved on the secure system in it's original compressed
format.

### 3. Installation

On the secure computing platform, once the archive has been transferred, you can then install the package.
The installation method differs depending on the filetype.

### 3.1 Install `.tar.gz` archives

Using an appropriate relative path for the archive, you can install it from source:

```R
# install the package directly from source
install.packages("./nutrientprofiler-v1.0.0.tar.gz", repos = NULL, type="source")
```


Change the suggested filename to suit your specific installation:

```R
# install the package directly from source
install.packages("./nutrientprofiler-VarEdits.tar.gz", repos = NULL, type="source")
```

### 3.2 Install `.zip` archives

In order to install the package from a `.zip` file, you need to use the `devtools` package:

```R
# install and load devtools
install.packages("devtools")
library(devtools)
devtools::install_local("./nutrientprofiler-v1.0.0.zip")
```
Again, chasnge the suggested filename to suit your specific installation.




<!-- 


The `nutrientprofiler` package aims to provide a series of R functions that 
implement UK Nutrient Profiling Model scoring that can be applied across product 
datasets.

This is broken up into 3 parts within the package: 

- Specific gravity conversions
- Scoring the different nutritional components of the NPM score
- Calculating the NPM score and NPM assessment logic

To start we load the package the with command below.
```{r setup}
library(nutrientprofiler)
```

We can also load some example data to help us with the following examples.
This data was created specifically for this package and so is potentially 
quite different to how your real world data might look.

```{r}
dim(cdrcdrinks)
cdrcdrinks[1, ]
```

## Specific gravity conversions

When trying to determine the NPM score the volume or weight of the product needs
to be adjusted to account for it's specific gravity. 

To adjust product weights or volumes for specific gravity we use the `SGConverter`
function. This is a high level function that is designed to operate on each row of 
a data.frame parsing multiple columns to determine how to calculate the adjusted 
specific gravity score.

```{r eval=FALSE, class.source="bg-warning"}
Warning: The `SGConverter` function has been specifically designed with an
existing dataset in mind and expects specific column names to work.
```

```{r}
# using dplyr
library(dplyr)
cdrcdrinks %>%
  rowwise() %>%
  mutate(out = SGConverter(pick(everything()))) %>%
  select(out)

# using base R
cdrcdrinks["sg"] <- unlist(lapply(seq_len(nrow(cdrcdrinks)), function(i) SGConverter(cdrcdrinks[i, ])))
```

The below figure attempts to outline the hierarchy of function calls that `SGConverter` 
initiates. The logic for determining how to adjust values for specific gravity is 
complicated by the potential options around whether a drink is ready-to-drink,
a powdered preparation, or a cordial, within both the powdered and cordial options
additional consideration must be given to the preparation instructions that are provided.
In the below figure each function is named in each node and each column name the function 
uses to dispatch to underlying functions is specified in single quotes. If a function returns
a value it is marked with an empty diamond.

```{r diagram-mermaid, fig.align='center', fig.cap="SGConverter logic", fig.dim=c(10, 6), echo=FALSE}
library(DiagrammeR)
DiagrammeR::mermaid("
graph TB;
  1(SGConverter<br>'product_type') --> 2(sg_food_converter<br>'weight_g');
  1 --> 3(sg_drink_converter<br>'drink_format');
  3 --> 4(sg_ready_drink_converter<br>'drink_type');
  3 --> 5(sg_powd_drink_converter<br>'nutrition_info');
  3 --> 6(sg_cord_drink_converter<br>'nutrition_info');
  6 --> 7{ };
  2 --> 10{ };
  2 --> 11(sg_liquidfood_converter<br>'food_type');
  11 --> 13{ };
  4 --> 14{ };
  5 --> 16{ };
")
```

## Nutrient profile model scoring

The next part of this package is a series of functions for handling Nutrient
Profile Model scoring. These specifically look at functions for adjusting units
into the expected unit as documented for the [Nutrient Profile
Model](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/216094/dh_123492.pdf),
a generic scoring function for returning a score for a value given a number of
thresholds, a scoring dispatcher function that determines what adjuster and
scoring thresholds to use for a given value and a high level `NPMScore` wrapper
function that shows the logic for calculating the NPM score across multiple
nutritional groups for a data.frame of data.

A quick example of running `NPMScore` across a single row of data is shown below:
```{r}
NPMScore(cdrcdrinks[1,], sg_adjusted_label="sg")
```

An example of building a tidyverse pipeline to calculate scores across all rows
of a data.frame using `NPMScore` and `SGConverter` is shown below: 

```{r}
library(tidyr)

cdrcdrinks %>% 
  rowwise() %>% 
  mutate( sg = SGConverter(pick(everything()))) %>% 
  mutate(test = NPMScore(pick(everything()), sg_adjusted_label="sg")) %>% 
  unnest(test) %>% select(energy_score, sugar_score, salt_score, fvn_score,
  protein_score, satfat_score, fibre_score)
```

## Nutrient Profile Model assessment

After calculating the scores for specific nutrients for our product next we need
to perform the actual Nutrient Profile Model assessment. This involves combining
individual nutrient scores to calculate an A score and a C score and use these
compound to calculate a total Nutrient Profile Model score and assess this to
determine a pass or fail.

This package implements this logic in a high-level wrapper function called
`NPMAssess` which operates on a row of a data.frame. It expects the columns
generated in the previous `NPMScore` step to allow it to calculate the A score
and C score and can be used as follows:

```{r}
# create NPM_score data.frame from NPMScore
# using the specific gravity `sg` column created above
npm_scores <- do.call(
        "rbind",
        lapply(
            seq_len(nrow(cdrcdrinks)),
            function(i) NPMScore(cdrcdrinks[i, ], sg_adjusted_label = "sg")
        )
    )

# append NPM Score columns to original data
combo_df <- cbind(cdrcdrinks, npm_scores)

# test NPMAssess on this data
NPMAssess(combo_df[1, ])
```

We can also use tidyverse functions to build an entire pipeline for running
Nutrient Profile Model assessments.

```{r}
# using tidyr
library(tidyr)

cdrcdrinks %>% 
  rowwise() %>% 
  mutate( sg = SGConverter(pick(everything()))) %>% 
  mutate(test = NPMScore(pick(everything()), sg_adjusted_label="sg")) %>% 
  unnest(test) %>% 
  rowwise() %>%
  mutate(assess = NPMAssess(pick(everything()))) %>%
  unnest(assess) %>%
  select(product_type, NPM_score, NPM_assessment)
```


The building blocks of `NPMAssess` are explained in more detail in the [Nutrient
Profile Model Assessment vignette](./nutrientprofile_assessment.Rmd) -->
