---
title: "General guidance on using this package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{General guidance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
library(nutrientprofiler)
example_data <- read.csv(text = "
name,brand,product_category,product_type,food_type,drink_format,drink_type,nutrition_info,energy_measurement_kj,energy_measurement_kcal,sugar_measurement_g,satfat_measurement_g,salt_measurement_g,sodium_measurement_mg,fibre_measurement_nsp,fibre_measurement_aoac,protein_measurement_g,fvn_measurement_percent,weight_g,volume_ml,volume_water_ml
lembas,,,Food,,,,,266,,50,3,,0.6,3,,7,0,100,,
zeno's icecream,,,Food,Ice cream,,,,,24,21,11,0.08,,,0.7,3.5,0,,100,
mystic rush,,,Drink,,Ready,Carbted/juice drink,,,194,11,0,,100,,0,0,0,,100,
delta ringer drink,,,Drink,,Powdered,,Preparation instructions given,188,,15,0,,100,,0,0.5,3,25,,100
welter water,,,Drink,,Cordial,,As consumed,,205,19,0,0.1,,,0,0.1,6,,100,
janus's drink,,,Food,,,,,,24,21,11,0.08,,,0.7,3.5,0,,100,
beta ringer drink,,,Drink,,Powder,,As consumed,188,,15,0,,100,,0,0.5,3,,50,
zeta ringer drink,,,Drink,,Powdered,,Prep. instructions not given,188,,15,0,,100,,0,0.5,3,25,,
heavyweight water,,,Drink,,Cordial,,Preparation instructions given,,205,19,0,0.1,,,0,0.1,6,,20,100
bantam water,,,Drink,,Cordial,,Preparation instructions not given,,205,19,0,0.1,,,0,0.1,6,,100,")
```

For this guidance, we are going to use an example dataset called `example_data`. Please see the detailed documentation on what parameter names and values are expected or required, and to see the workflows explaining how to load in data.

This example data has a few typos in certain categories. This documentations steps through one example workflow for filtering and checking your input data before using the `nutrientprofiler` library to analyse it.

```R
# load in required libraries
library(nutrientprofiler)
# read in the data
example_data <- read.csv("data/example_data.csv")
```

## Pre-processing your data

While there are an array of functions to check parameter names and missing parameters, it is useful to be able to manually check your input data with base R or a library like tidyr or dplyr additionally.

```{r}
# Print out your data column names
names(example_data)
```
