---
title: "Handling input data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling input data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Input data is likely to be messy and to not contain the required column names for `nutrientprofiler`.
This vignette provides some advice and guidelines on how to pre-process your data to allow analysis
and assessment with `nutrientprofiler`.

You can see what parameters are expected in more detail in the [Example dataset reference](../reference/cdrcdrinks.html).

```{r setup}
library(nutrientprofiler)

irregular_data <- cdrcdrinks
names(irregular_data)[names(irregular_data) == "fvn_measurement_percent"] <- "fruit_nutrition_percent"
```

## Loading an example template

When possible, it is a good idea to ensure your data is in the required format before yourload it in to R. This can be achieved by using a template.

The `nutrientprofiler` package offers an example dataset, `cdrcdrinks`. This can be saved out as an
example csv file and then edited and loaded with real data:

```R
# Save an example csv file to your current working directory
write.csv(cdrcdrinks, "template_full.csv", row.names = FALSE)
```

Alternatively, just the headers can be saved, providing a blank csv file which can be opened in
Excel or your editor of choice:

```R
# Save an example csv file to your current working directory
write.csv(t(names(cdrcdrinks)), "template_headings.csv", row.names = FALSE)
```

## Checking column names and renaming parameters

If raw data has been acquired with different column names than required, a reproducible
workflow can be set up to rename these parameters as required.

The first step is to check the existing column names:

```{r}
inputDataCheck(cdrcdrinks)
```

If the dataframe contains names as expected, you'll get a message telling you to proceed with analysis,
as with the example dataset above.

However, lets say we have loaded in a dataset from csv `irregular_data` that does not contain the
exact columns required:

```{r}
inputDataCheck(irregular_data)
```

This provides you with the name of the column(s) that do not match an expected column name
(`"fruit_nutrition_percent"` in this example), and provides you with a prompt to use the
`parameterRename()` function.

You can copy this hint and replace `<INSERT DATA COLUMN NAME HERE>` with the data that matches
the listed `missing_column_name`:

```{r}
irregular_data <- parameterRename(missing_column_name = 'fvn_measurement_percent', associated_data_column = 'fruit_nutrition_percent', data_frame = irregular_data)
```

You can now check again if the required parameters are available.

```{r}
inputDataCheck(irregular_data)
```

