---
title: "Nutrient Profile Model Assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nutrient Profile Model Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For assessing individual nutrient scores of a product within the Nutrient
Profile Model we have to complete the following steps:

- Calculate compound A and C scores 
- Calculate a total Nutrient Profile Model score using A and C scores 
taking into account Fibre and fruit, vegetables and nut scores 
- Determine whether the product passes or fails with thresholds depending on
whether the product is a food or drink

The `nutrientprofiler` package includes a high level wrapper function
`NPMAssess` for handling this row-by-row on a data.frame with columns generated
from `NPMScore`. Below we will explore the component functions of `NPMAssess` to
understand how this function operates.

```{r setup}
library(nutrientprofiler)
```

## A and C score calculators

The first component of `NPMAssess` is the step to calculate the A and C scores.
This involves two functions: `A_scorer` and `C_scorer` that accept a number of
arguments corresponding to the appropriate nutrient score and sum them.

The `A_scorer` takes 4 arguments: `energy_score`, `sugar_score`, `salt_score` and
`fat_score`. These are scores calculated in the previous `NPMScore` step and are
summed in the `A_scorer` function to return a compound A score. The `C_scorer`
takes 3 arguments: `fvn_score` (fruit, vegetable and nut score), `protein_score`
and `fibre_score` and calculates the sum of these to return a C score.

