---
title: "Introduction to nutrientprofiler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to nutrientprofiler}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script src="https://cdn.jsdelivr.net/npm/mermaid/dist/mermaid.min.js"></script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `nutrientprofiler` package aims to provide a series of R functions that 
implement UK Nutrient Profiling Model scoring that can be applied across product 
datasets.

This is broken up into 3 parts within the package: 

- Specific gravity conversions
- Scoring the different nutritional components of the NPM score
- Calculating the NPM score and NPM assessment logic

To start we load the package the with command below.
```{r setup}
library(nutrientprofiler)
```

We can also load some example data to help us with the following examples.
This data was created specifically for this package and so is potentially 
quite different to how your real world data might look.

```{r}
dim(cdrcdrinks)
cdrcdrinks[1, ]
```

## Specific gravity conversions

When trying to determine the NPM score the volume or weight of the product needs
to be adjusted to account for it's specific gravity. 

To adjust product weights or volumes for specific gravity we use the `SGConverter`
function. This is a high level function that is designed to operate on each row of 
a data.frame parsing multiple columns to determine how to calculate the adjusted 
specific gravity score.

```{r eval=FALSE, class.source="bg-warning"}
Warning: The `SGConverter` function has been specifically designed with an
existing dataset in mind and expects specific column names to work.
```

```{r}
# using dplyr
library(dplyr)
cdrcdrinks %>%
  rowwise() %>%
  mutate(out = SGConverter(pick(everything()))) %>%
  select(out)

# using base R
cdrcdrinks["sg"] <- unlist(lapply(seq_len(nrow(cdrcdrinks)), function(i) SGConverter(cdrcdrinks[i, ])))
```

The below figure attempts to outline the hierarchy of function calls that `SGConverter` 
initiates. The logic for determining how to adjust values for specific gravity is 
complicated by the potential options around whether a drink is ready-to-drink,
a powdered preparation, or a cordial, within both the powdered and cordial options
additional consideration must be given to the preparation instructions that are provided.
In the below figure each function is named in each node and each column name the function 
uses to dispatch to underlying functions is specified in single quotes. If a function returns
a value it is marked with an empty diamond.

```{r diagram-mermaid, fig.align='center', fig.cap="SGConverter logic", fig.dim=c(10, 6), echo=FALSE}
library(DiagrammeR)
DiagrammeR::mermaid("
graph TB;
  1(SGConverter<br>'product_type') --> 2(sg_food_converter<br>'weight_g');
  1 --> 3(sg_drink_converter<br>'drink_format');
  3 --> 4(sg_ready_drink_converter<br>'drink_type');
  3 --> 5(sg_powd_drink_converter<br>'nutrition_info');
  3 --> 6(sg_cord_drink_converter<br>'nutrition_info');
  6 --> 7{ };
  2 --> 10{ };
  2 --> 11(sg_liquidfood_converter<br>'food_type');
  11 --> 13{ };
  4 --> 14{ };
  5 --> 16{ };
")
```

## Nutrient profile model scoring

The next part of this package is a series of functions for handling Nutrient
Profile Model scoring. These specifically look at functions for adjusting units
into the expected unit as documented for the [Nutrient Profile
Model](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/216094/dh_123492.pdf),
a generic scoring function for returning a score for a value given a number of
thresholds, a scoring dispatcher function that determines what adjuster and
scoring thresholds to use for a given value and a high level `NPMScore` wrapper
function that shows the logic for calculating the NPM score across multiple
nutritional groups for a data.frame of data.

A quick example of running `NPMScore` across a single row of data is shown below:
```{r}
NPMScore(cdrcdrinks[1,], sg_adjusted_label="sg")
```

An example of building a tidyverse pipeline to calculate scores across all rows
of a data.frame using `NPMScore` and `SGConverter` is shown below: 

```{r}
library(tidyr)

cdrcdrinks %>% 
  rowwise() %>% 
  mutate( sg = SGConverter(pick(everything()))) %>% 
  mutate(test = NPMScore(pick(everything()), sg_adjusted_label="sg")) %>% 
  unnest(test) %>% select(energy_score, sugar_score, salt_score, fvn_score,
  protein_score, satfat_score, fibre_score)
```

## Nutrient Profile Model assessment

After calculating the scores for specific nutrients for our product next we need
to perform the actual Nutrient Profile Model assessment. This involves combining
individual nutrient scores to calculate an A score and a C score and use these
compound to calculate a total Nutrient Profile Model score and assess this to
determine a pass or fail.

This package implements this logic in a high-level wrapper function called
`NPMAssess` which operates on a row of a data.frame. It expects the columns
generated in the previous `NPMScore` step to allow it to calculate the A score
and C score and can be used as follows:

```{r}
# create NPM_score data.frame from NPMScore
# using the specific gravity `sg` column created above
npm_scores <- do.call(
        "rbind",
        lapply(
            seq_len(nrow(cdrcdrinks)),
            function(i) NPMScore(cdrcdrinks[i, ], sg_adjusted_label = "sg")
        )
    )

# append NPM Score columns to original data
combo_df <- cbind(cdrcdrinks, npm_scores)

# test NPMAssess on this data
NPMAssess(combo_df[1, ])
```

We can also use tidyverse functions to build an entire pipeline for running
Nutrient Profile Model assessments.

```{r}
# using tidyr
library(tidyr)

cdrcdrinks %>% 
  rowwise() %>% 
  mutate( sg = SGConverter(pick(everything()))) %>% 
  mutate(test = NPMScore(pick(everything()), sg_adjusted_label="sg")) %>% 
  unnest(test) %>% 
  rowwise() %>%
  mutate(assess = NPMAssess(pick(everything()))) %>%
  unnest(assess) %>%
  select(product_type, NPM_score, NPM_assessment)
```


The building blocks of `NPMAssess` are explained in more detail in the [Nutrient
Profile Model Assessment vignette](./nutrientprofile_assessment.html)
