---
title: "Nutrient Profile Model Assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nutrient Profile Model Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For assessing individual nutrient scores of a product within the Nutrient
Profile Model we have to complete the following steps:

- Calculate compound A and C scores 
- Calculate a total Nutrient Profile Model score using A and C scores 
taking into account Fibre and fruit, vegetables and nut scores 
- Determine whether the product passes or fails with thresholds depending on
whether the product is a food or drink

The `nutrientprofiler` package includes a high level wrapper function
`NPMAssess` for handling this row-by-row on a data.frame with columns generated
from `NPMScore`. Below we will explore the component functions of `NPMAssess` to
understand how this function operates.

```{r setup}
library(nutrientprofiler)
```
