---
title: "Nutrient Profile Model Scoring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nutrient Profile Model Scoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To determine the individual nutritional element scores that can be used in a
Nutrient Profile Model there are two crucial steps:

- Adjust the nutritional value based on the specific gravity adjusted weight and
any unit transformations required 
- Score the adjusted nutritional value based on a set of thresholds

Below we outline the underlying logic for implementing both these steps in the
`nutrientprofiler` package.

```{r setup}
library(nutrientprofiler)
```

## Adjuster functions 

Nutritional values need to be adjusted before scoring to ensure:

- The nutritional information is adjusted based on their nutrient denisty which
is standardised to 100g of the product rather than the total amount to enable 
better comparisons between products
- They are converted into the appropriate unit for scoring thresholds

All adjuster functions are built ontop of the `generic_adjuster` function. 

```{r}
generic_adjuster(10, 15)
```

This takes a `value` argument for the amount of a nutritional category and
divides it by the second argument `adjusted_weight` which corresponds to the
specific gravity adjusted weight/volume for the product as determined using the
specific gravity functions above. The output of this is multiplied by 100 to
give the final adjusted value.

### Energy and salt adjustments

This is enough for adjusting most nutritional values but for both energy and
salt we need additional steps to properly adjust the value for the appropriate
units. This is because for energy nutritional information can be in either
*kcal* or *kJ* and for salt can either be listed as a sodium measurement in *mg*
or a salt measurement in *g*. The scoring thresholds for the Nutrient Profile
Model are specified for *kJ* for energy and *mg* of sodium for salt so
additional steps are needed to properly adjust these values if present in the
alternate unit.

This is handled by specific adjuster functions for energy and salt that also
take an `adjuster_type` argument to specify the units of the passed value.

```{r}
# for energy in kj
energy_value_adjuster(10, 80, adjuster_type="kj")

# for energy in kcal
energy_value_adjuster(60, 300, adjuster_type="kj")

# for salt in grams
salt_adjuster(3, 100, adjuster_type="salt")

# for salt in mg of sodium
salt_adjuster(60, 100, adjuster_type="sodium")
```

## Scoring function 

The adjuster functions return an adjusted value for nutritional information to
ensure we can calculate the NPM Score based on 100g of the product. We use this
adjusted value in the subsequent step for actually calculating a score for each
of our nutrient groups. This is performed using the `NPM_score_function` which
acts as a dispatcher function that takes `value` and `type` arguments to
determine which adjustment function to call and which scoring thresholds to use.
This function also takes optional additional arguments such as specifying the
`adjusted_weight` that can be passed into the adjuster functions and the
`adjuster_type` argument  for some types such as `energy` and `salt`.

```{r}
# calculate a score for a protein value
NPM_score_function(80, adjusted_weight=100, "protein")
```

Under the hood this function is calling the `generic_adjuster` from above to
adjust the nutritional value using the `adjusted_weight`, next it passes this
adjusted value to the `scoring_function`. The `scoring_function` determines the
score of a value based on provided score thresholds.

```{r}
# create thesholds vector from largest to smallest
THRESHOLDS <- c(3, 2, 1)

# score the value 1 against the provided thresholds
# returns 0
scoring_function(1, THRESHOLDS)

# returns 2
scoring_function(2.1, THRESHOLDS)
```

The `scoring_function` behaves by taking a vector of thresholds in descending
order. First it sets the score to be the total number of thresholds available (i.e. 3 in the above
example). It takes the value we have provided and checks if this value is
greater than each item in the thresholds vector, if it is greater than the item
from thresholds it returns the score corresponding to that item. If the value
isn't greater than the item then the score, set earlier, is deducted 1. The
continues for each item in the thresholds until the only option left is our
value not being greater than the final item in our thresholds leading to a score
of 0.

### Fruit, vegetables and nuts scoring 

The one exception to this scoring function is for fruit, vegetables and nuts.
This nutritional information is presented as a percentage and the scoring for
this specified in the Nutrient Profile Model doesn't score this in a continuous
manner so this package includes a specific `fruit_veg_nut_scorer`. This function
takes value and returns a score (either 0, 1, 2 or 5) depending on if the value
meets a specific percentage threshold.

```{r}
# scores a 5
fruit_veg_nut_scorer(81)

# scores a 2
fruit_veg_nut_scorer(70)

# scores a 1
fruit_veg_nut_scorer(45)

# scores a 0
fruit_veg_nut_scorer(20)
```