---
title: "Specific Gravity Adjustment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Specific Gravity Adjustment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When attempting to calculate a Nutrient Profile Model volumes and masses need to
be adjusted by the specific gravity of the product to ensure a correct reading.
This is managed directly by the `SGConverter` high level function that operates
row-by-row on a data.frame with a series of expected columns. This function is
composed of a series of different dispatcher function which will be outlined
below (see figure for hierarchy).

```{r diagram-mermaid, fig.align='center', fig.cap="SGConverter logic", fig.dim=c(10, 6), echo=FALSE}
library(DiagrammeR)
DiagrammeR::mermaid("
graph TB;
  1(SGConverter<br>'product_type') --> 2(sg_food_converter<br>'weight_g');
  1 --> 3(sg_drink_converter<br>'drink_format');
  3 --> 4(sg_ready_drink_converter<br>'drink_type');
  3 --> 5(sg_powd_drink_converter<br>'nutrition_info');
  3 --> 6(sg_cord_drink_converter<br>'nutrition_info');
  6 --> 7{ };
  2 --> 10{ };
  2 --> 11(sg_liquidfood_converter<br>'food_type');
  11 --> 13{ };
  4 --> 14{ };
  5 --> 16{ };
")
```

```{r setup}
library(nutrientprofiler)
```

The central logic for specific gravity conversion is multiplying the
volume/weight of a product by a specific gravity multiplier. This is distilled
in he `generic_specific_gravity` function. 

```{r}
generic_specific_gravity(100, 1.01)
```

In practise this package uses the `SGtab` named vector which contains a series
of named products and their associated specific gravity. This variable is not
directly accessible to users but is used by other specific gravity functions to
determine the appropriate specific gravity adjustments.

## Breaking down `SGConverter`

`SGConverter` works by taking the row of a data.frame as it's argument. It
checks that row for a column called `product_type`. This column is used to
define whether the product is either a `"food"` or `"drink"`. The value in this
column is used by `SGConverter` to determine whether to pass the row to either
an `sg_food_converter` function or a `sg_drink_converter` function.

### Calculating specific gravity for food

If `SGConverter` is passed a `product_type` that is `"food"` it passes the row
to `sg_food_converter`. This function does a simple test of whether the row
contains a column called `weight_g`, if this column doesn't contain `NA` it
returns the value contained in `weight_g` as we don't have to adjust this for a
specific gravity.
If the `weight_g` column does contain `NA` the row is dispatches to the
`sg_liquidfood_converter` function. The `sg_liquidfood_converter` checks for a
`food_type` column which contains a type value that should match a product
included in our `SGtab` vector. The `sg_liquidfood_converter` checks if the
`food_type` column is not an empty string, if this is true it calculates the
specific gravity adjusted volume by getting the value from the `volume_ml`
column in the row and multiplying it by the specific gravity value in the
`SGtab` which is indexed out of the vector using the `food_type` column value
from the passed row.

### Calculating specific gravity for drinks 

The logic for calculating specific gravity for drinks is slightly more
complicated. If the `SGConverter` function found that `"drink"` was specified in
the `product_type` column it dispatches the row to the `sg_drink_converter`
function. This checks the row for a column called `drink_format` that specifies
whether the drink is either: ready to drink, a cordial or a powdered drink
(`"ready"`, `"cordial"`, `"powdered"`). For each of these options the
`sg_drink_converter` dispatches the row to an additional function.

For `"ready"` it dispatches to `sg_ready_drink_converter`. This function checks
if the row contains a `drink_type` column and if that row contains an empty
string. If it does not contain an empty string it calls the
`generic_specific_gravity` function passing the `volume_ml` column value and the
specific gravity multiplier indexed from `SGtab` based on the value in
`drink_type`. If `drink_type` does contain an empty string it just returns the
`volume_ml` column value.

For `"cordial"` it dispatches to `sg_cord_drink_converter`. This function checks
the row for a `nutrition_info` column to help determine how it should proceed.
This column contains information about how the drink has been consumed, whether
it has been directly consumed, following some preparation instructions that
dilute the cordial or if no instructions were provided (`"as consumed"`,
`"preparation instructions given"`, `"preparation instructions not given"`). 
If `nutrition_info` is `"as consumed"` the `volume_ml` column value and the
specific gravity multipler from the `SGtab` for 
`"Cordial/squash ready to drink"` is passed to the `generic_specific_gravity`
function. 
For `"preparation instructions given"` the expectation is that there is an
additional column in the row `volume_ml_water` that specifies the volume of
water used to dilute the cordial. If `"preparation instructions given"` is found
in `nutrition_info` column it computes the sum of `volume_ml` and
`volume_ml_water` columns and passes those to `generic_specific_gravity` along
with the specific gravity multipler from the `SGtab` for  `"Cordial/squash ready to drink"`.
For `"preparation instructions not given"` the `volume_ml` column value is
passed to `generic_specific_gravity` function along with the specific gravity
multiplier indexed from `SGtab` for `"Cordial/squash undiluted"`.

If the `drink_format` column value is `"powdered"` it dispatches to the
`sg_powd_drink_converter` function. This again checks the row for values in the
`nutrition_info` column to help determine how to compute the specific gravity
adjustment using the same 3 potential values defined above in the `cordial`
section. 
If `nutrition_info` is `"as consumed"` the `volume_ml` column value and the
specific gravity multipler from the `SGtab` for 
`"Cordial/squash ready to drink"` is passed to the `generic_specific_gravity`
function. 
For `"preparation instructions given"` the expectation is that there is an
additional column in the row `weight_g` that specifies the weight in grams of
the powdered drink used. If `"preparation instructions given"` is found
in `nutrition_info` column it computes the sum of `weight_g` and
`volume_ml_water` columns and passes those to `generic_specific_gravity` along
with the specific gravity multipler from the `SGtab` for  `"Cordial/squash ready to drink"`.
For `"preparation instructions not given"` the `weight_g` column value is
returned unadjusted.

## Summary

Overall most of the functions in `SGConverter` are quite constrictive. They
expect particular column names to work and won't just work with any data.
However, the building blocks to all these functions is the
`generic_specific_gravity` which can be used to develop your own functions for
building a specific gravity conversion pipeline. The existing steps shown here
should serve as a template to build your own specific gravity conversion pipeline.